---
title: "Visualize Community-based Features"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(ggthemes)
require(igraph)

wide_to_long <- function(df, id_var, key_var, val_var, sort_by){
    df %>% t() %>% as.data.frame() %>% 
        rownames_to_column(id_var) %>% 
        arrange(!!sort_by) %>% 
        gather(!!key_var, !!val_var, -(!!id_var)) 
}

poster_theme <- theme(plot.title = element_text(size = 24, face = "bold"), 
                      plot.subtitle =  element_text(size = 10), 
                      axis.text = element_text(size = 15), 
                      axis.title=element_text(size=14,face="bold")) +  
    theme(rect = element_rect(fill = "transparent")) 
poster_scale_color <- scale_color_brewer(palette="Dark2") 
    

cols <- c("Rstudio-related", "base", "Image Plotting",
          "RCpp", "GPS and Geography", "Machine learning",
          # "public health and Statistics", # error
          "Text Analysis", "Social Network Analysis", "Graphics", 
          "Graph", "Genetics", "Finance", 
          "Insurance and Actuary", "Numerical Optimization", "Sparse Matrix", 
          "Java", "Time, Date, and Money", "Neuronal Science")

community_ids <- list(15, 9, 4, 
                      60, 14, 35, 
                      # 7,# error, shall be 1
                      36, 25, 39, 
                      23, 19, 31, 
                      8, 64, 73, 
                      18, 20, 120)

cran_graph <- read_rds("cran_graph.RDS")
cran_wc <- read_rds("cran_community_20190518.RDS")
cran_event <- evcent(cran_graph, direct = TRUE)
get_key_members <- function(target_community){
    idx <- which(cols == target_community)
    target_community_id <- community_ids[idx]
    key_members <- tibble(pkg = V(cran_graph)$name, 
                          comm = cran_wc$membership, 
                          evcent = cran_event$vector) %>% 
        filter(comm == target_community_id) %>% 
        arrange(desc(evcent)) %>% 
        head(10) %>% 
        pull(pkg)
    return(key_members)    
}

size_community <- tibble(pkg = V(cran_graph), comm = cran_wc$membership) %>% group_by(comm) %>% summarize(n = n()) %>% arrange(desc(n))
get_size_community <- function(target_community){
    idx <- which(cols == target_community)
    target_community_id <- community_ids[idx]
    size_community %>% filter(comm == target_community_id) %>% pull(n)
}

# naming
df_naming_ratio_total <- readRDS('community_df_naming_ratio_total.RDS') %>% select(-`public health and Statistics`)
colnames(df_naming_ratio_total) <- cols
rownames(df_naming_ratio_total) <- c("alllower", "ALLUPPER", "UpperCamel", "lowerCamel", "lower_snake", "dotted.func", "other")

# features
df_ratio_total <- readRDS('community_df_ratio_total.RDS') %>% as.data.frame() %>% select(-`public health and Statistics`)
colnames(df_ratio_total) <- cols
# rownames(df_ratio_total) <- c("")
```

## community

#### what's the size of each community? Also the top 10 package members?
```{r, echo=FALSE}
size_community
```


## function naming convention
```{r, include=FALSE}
levels_naming <- c("dotted.func", "ALLUPPER", "UpperCamel", "other", "alllower", "lowerCamel", "lower_snake")
df_naming_ratio_long <- df_naming_ratio_total %>% 
    wide_to_long(id_var = 'community', key_var = 'naming', val_var = 'ratio', sort_by = quo(lower_snake)) %>% 
    mutate(percentage = ratio*100, 
           community=factor(community, levels = unique(community)),
           naming=factor(naming, levels = levels_naming)) 
```

#### overall

#### what's most popular naming convention within community?

- Summary
    - Anomaly Detection is a big fan of lowerCamel, 78%, of great harmony within community. (But suffer from open curly problem most)
    - Java has its own styles => other
    - rstudio is the main advocate of lower_snake style, consistent with the overall trend across years above
    - dotted.fun gains great support from social network analysis
    - Finance, for some reason, uses UpperCamel, which is not common in other communities
    - ALLUPPER is not an abandoned naming
    - dotted.func, only proportional

- Historical reason for them?

```{r, echo=FALSE}
idxmax <- function(vals){
    idx <- which(vals == max(vals))
    return(names(vals)[idx])
}
df_pop <- df_naming_ratio_total %>% t() %>% as.data.frame()
df_pop_summary <- data.frame(community = df_pop %>% apply(2, idxmax), ratio = df_pop %>% apply(2, max)) 
df_pop_summary %>% rownames_to_column(var = "naming") %>% arrange(-ratio) %>% select(community, naming, ratio)

```

- main stream community
    - rstudio
- Community that is different from others
    - anomaly detection: lowerCamel
    - Finance: UpperCamel
    - Java: other
    - Public health: little snake/alllower
```{r naming, echo=FALSE, fig.height = 6, fig.width = 9}
g_naming <- ggplot(aes(y = percentage, x = community, fill = naming), data = df_naming_ratio_long) + 
    geom_bar(stat="identity") + 
    labs(x = "", y = "%") + 
    theme(legend.title = element_blank()) +
    coord_flip() + 
    poster_theme + poster_scale_color

print(g_naming)
ggsave("visualization_community/naming_among_community.png", plot = g_naming, width = 9, height = 6, units = "in", bg = "transparent")
```

#### within each community
```{r, echo=FALSE, fig.height = 4, fig.width = 6}

wrapper <- function(x, ...) 
{
  paste(strwrap(x, ...), collapse = "\n")
}

get_key_members_subtitle <- function(target_community){
    members <- get_key_members(target_community) 
    top10_members <- paste(members, collapse = ", ")
    total_members <- get_size_community(target_community)
    subtitle <- sprintf("TOP 10 / %s: %s", total_members, top10_members )
    wrapper(subtitle, width=60)
}

gen_fig_naming_dist <- function(target_community){
    df_plot <- df_naming_ratio_long %>% filter(community == target_community, 
                                               naming %in% c("lower_snake", "lowerCamel", "UpperCamel", "dotted.func"))
    g <- ggplot(aes(y = percentage, x = naming, fill=naming), data = df_plot) + 
        geom_bar(stat="identity") + ylim(0, 100) +         
        labs(title = target_community, subtitle = get_key_members_subtitle(target_community), x = "", y = "") + 
        theme_hc() + theme(legend.position = "none", 
                           panel.grid.major.y = element_blank(), 
                           axis.text.x = element_blank(), 
                           axis.ticks.x = element_blank()) +
        coord_flip() +
        geom_text(aes(label=sprintf("%s %s", round(percentage, 2), "%")), position = position_dodge(width=0.9), hjust = -0.25) + 
        poster_theme + poster_scale_color
    return(g)
}

communities <- as.character(unique(df_naming_ratio_long$community))
for (comm in communities){
    print(comm)
    g <- gen_fig_naming_dist(comm)
    print(g)
    filename <- sprintf("visualization_community/naming_in_comm_%s.png", comm)
    ggsave(filename, plot = g, width = 6, height = 4, units = "in", bg = "transparent")
}
```




## features 

#### distance
```{r, echo=FALSE}
avg_feature_error <- df_ratio_total %>% rowMeans()
get_distance <- function(target_community){
    feature_error <- df_ratio_total[, target_community]
    dist <- dist(matrix(c(feature_error, avg_feature_error), nrow = 2))[1]
    return(dist)
}

get_size_community_v <- function(communities){
  sapply(communities, get_size_community)
}

communities <- as.character(unique(df_naming_ratio_long$community))
dist_comm <- sapply(communities, get_distance)
z_dist_comm <- (dist_comm - mean(dist_comm))/sd(dist_comm)
df_dist <- bind_cols(community = names(dist_comm), dist = dist_comm, zdist = z_dist_comm) %>%
  mutate(size = get_size_community_v(community)) %>%
  arrange(dist)
df_dist

```

```{r, echo=FALSE, fig.height=6, fig.width=12} 
wt <- df_dist$size/sum(df_dist$size)
mu_dist <- weighted.mean(df_dist$dist, wt)
sigma_dist <- sum(wt * (df_dist$dist - mu_dist)^2) %>% sqrt()

df_dist_plot <- df_dist %>% 
  add_row(community = "average", dist = 0, zdist = 0, .before = 1) %>% 
   mutate(community = factor(community, levels = unique(community)))

g_dist <- df_dist_plot %>% ggplot(aes(x = community, y = dist)) + 
  geom_point() + 
  geom_segment(aes(x=community, xend = community, y = 0, yend = dist)) +
  labs(x = "", y = "distance") + 
  geom_hline(yintercept = mu_dist, color = 'red', alpha = 0.4) + 
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = mu_dist + sigma_dist * 1, ymax = mu_dist + sigma_dist * 2, alpha = 0.1, fill = "red") + 
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = mu_dist + sigma_dist * 2, ymax = mu_dist + sigma_dist * 3, alpha = 0.2, fill = "red") +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = mu_dist + sigma_dist * 3, ymax = Inf, alpha = 0.4, fill = "red") +
  geom_text(aes(label = round(dist, 2)), position = position_dodge(width=0.9), hjust = -0.40) + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  poster_theme + poster_scale_color + 
  coord_flip()

print(g_dist)

ggsave("visualization_community/feature_distance.png", plot = g_dist, width = 12, height = 6, units = "in", bg = "transparent")
```


- settle down: most-agreed
    - fx_close_curly
    - fx_semi
    - fx_t_f
    - fx_singleq
    
- rstudio does not good at
    - fx_integer (nr. 1)
    - fx_inflix
    - fx_assign
    - fx_sigleq
    
- lowest variation inside community
    - time, date, and money
    
- everyone need to improve
    - fx_integer
    - fx_infix

```{r features, echo=FALSE, fig.height = 12, fig.width = 12}
levels_feature <- apply(df_ratio_total, 1, max) %>% sort(decreasing = TRUE) %>% names()

df_ratio_long <- df_ratio_total %>% 
    wide_to_long(id_var = 'community', key_var = 'feature', val_var = 'ratio', sort_by = quo(fx_opencurly)) %>% 
    mutate(percentage = ratio*100, 
           community= factor(community, levels = unique(community)),
           feature = factor(feature, levels_feature)) 

g_features <- ggplot(aes(y = percentage, x = community), data = df_ratio_long) + 
    geom_bar(stat="identity", position="dodge") +
    labs(x = "") + 
    facet_grid(feature ~ ., switch="y") + 
    scale_y_continuous("percentage", position="right") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          strip.text.y=element_text(angle=180)) +
    poster_theme + poster_scale_color

print(g_features)
ggsave("visualization_community/features_among_community.png", plot = g_features, width = 10, height = 16, units = "in", bg = "transparent")
```

#### selected features

```{r, echo=FALSE, fig.height = 6, fig.width = 12}
levels_community <- unique(df_dist_plot$community) %>% as.character()

df_ratio_long_sub <- df_ratio_long %>% 
  filter(feature %in% c("fx_opencurly", "fx_integer", "fx_infix", "fx_assign")) 

df_ratio_sub_avg <- df_ratio_long_sub %>% 
  mutate(size = get_size_community_v(community), wt = size/sum(size)) %>% 
  group_by(feature) %>% summarise(ratio = weighted.mean(ratio, wt), 
                                  percentage = weighted.mean(percentage, wt)) %>%
  mutate(community = "average") %>% 
  select(community, feature, ratio, percentage)

df_ratio_long_sub_plot <- bind_rows(df_ratio_long_sub, df_ratio_sub_avg) %>% 
    mutate(color =  case_when(
        community == "average" ~ 1,
        community != "average" ~ 0)) %>% 
    mutate(community = factor(community, levels = levels_community), color = factor(color))

g_features_sub <- ggplot(aes(y = percentage, x = community, fill = color), data = df_ratio_long_sub_plot) +
  geom_bar(stat="identity", position="dodge") + 
  labs(x="", y = "") + 
  scale_x_discrete(position="top") + 
  scale_y_continuous(trans = 'reverse', limits = c(60, 0)) + 
  coord_flip() +
  facet_grid(. ~ feature) + 
  geom_text(aes(label=sprintf("%s%s", round(percentage, 0), "%")), position = position_dodge(width=0.9), hjust = 1.1) + 
  theme(legend.position = "none", 
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  poster_theme + poster_scale_color

print(g_features_sub)
ggsave("visualization_community/features_among_selected_community.png", plot = g_features_sub, width = 12, height = 6, units = "in", bg = "transparent")




```


